Increasing quality of self-recovering images using BPCS Steganography

### 2014-10-29 Paper/main.tex
Section 1 Completed

\title{Self recovering capabilities in images using the BPCS Steganography algorithm}

@@

\noindent 
Many current self-recovering image methods exist, however few of the current methods provide good quality images after embedding as well as after reconstruction.
This is due to an inherent trade off between the amount of information that can be embedded into an image and the quality of the image after embedding.
This dissertation proposes a novel method for increasing the quality of the image after embedding as well as the quality of the image after reconstruction in the self-recovering image problem using the BPCS Steganography algorithm for content embedding as well as employing a fragile watermark for detecting image modifications.
The proposed method achieves a good quality image after embedding as well as a good quality image after reconstruction when compared to some of the best currently available self-recovering image methods.

@@

\subsection{Introduction}
During communication images can by modified either intentionally or unintentionally.  These modifications can be small changes that does not change the meaning of the image, or bigger changes, for example the removal of key aspects from the image. 
Various techniques exist for authenticating images of which self-recovering is one technique. 
Self-recovering techniques attempt to preserve image content by embedding the image within itself to enable the reconstruction of the image if the image was tampered with. Self-recovering techniques use steganography, a technology used to embed information in digital media \cite{johnson1998exploring}, to hide the image content within the image.  Self-recovering techniques also use fragile watermarks which is a technology commonly used in image authentication, to determine which areas of an image was tampered with.

\hspace{0pt} \\
Most self-recovering techniques have a trade-off concerning the quality of the image after embedding and the quality of the image after recovery.
If the one's quality is increased the other's quality will be reduced.

@@

\subsection{Problem statement}
\label{problemStatement}
The proposed method aims to answer the following questions
\begin{itemize}
	\item Can a method be proposed that will achieve a high quality image after the self-embedding and restoration process as well as a reasonable tamper rate? How does such a system compare to currently available methods?
	\item Determining the quality of the resulting image should include, but are not limited to: 
	\begin{itemize}
		\item There should not be clear visual evidence that the image has been modified after embedding as well as after restoration
		\item The image should still carry it's original semantic meaning
	\end{itemize}
\end{itemize}

\subsection{Research methodology}
Firstly a literature review was conducted to determine the currently available methods for self-recovering images. 
After a shortcoming was identified the proposed method was designed and then implemented in the form of a Java software application in order to attempt to improve on the identified shortcoming. 
The results of the proposed method were then compared to the results of other relevant self-recovering image methods.

@@

\begin{itemize}
  \item Original Image: This refers to the original image without any modifications.
  \item Embedding: Embedding in the context of this dissertation refers to the process of storing digital information in an image's content in such a way that the original image is still distinguishable.
  \item Image after embedding: This refers to the the original image after the embedding process has been applied.
  \item Reconstruction: This refers to the process of using the content that was embedded into the image to repair the image after embedding if it has been modified.
  \item Image after reconstruction: This refers to the reconstructed version of the original image that was obtained by applying the reconstruction process to the image after embedding.
  \item LSB: This refers to the least significant bit in a binary string.
  \item MSB: This refers to the most significant bit/bits in a binary string.
  \item Section, Image section, Pixel section: This refers to a section of 8x8 pixels in an image.
  \item Bit plane: A bit plane is a 8x8 grid representation of a specific bit value for a specific pixel section. Each pixel has 24 bits which in total represents each colour, 8 bits per colour. These bits are the 8 binary bits that represent the numerical value of the colour. These 8 bits are separated from each other and are each stored in a 8x8 grid in the same location as that of the pixel in the pixel section. This process is repeated for each pixel in the pixel section and the result is 8, 8x8 grids of binary numbers. Each one of these 8 grids represent a single bit plane. 
  \item Block: This refers to an individual bit plane of a specific section of the image. There are 8 such blocks in each section since each pixel is represented by 8 bits for each colour.
\end{itemize}

\subsection{Chapter layout}
\begin{itemize}
\item Chapter 1 serves as an introduction to the concept of self-recovering images
\item Chapter 2 discusses self-recovering images and their different components
\item Chapter 3 discusses currently available self-recovering image schemes, comparing their strengths and weaknesses
\item Chapter 4 provides an overview of the system proposed in this dissertation
\item Chapter 5 discusses the results achieved during experimentation with the proposed system.
\item Chapter 6 Provides a conclusion, determining questions stated in section \ref{problemStatement} have been answered
\end{itemize}

\subsection{Conclusion}
This chapter discussed the purpose behind self-recovering images as well as the aim of this research and the proposed method. 
The research methodology followed was described. A list of terminologies used in this dissertation was also given.
The next chapter will give a more detailed overview of all the components that make up self-recovering images.

@@

\hspace{0pt} \\
The PSNR of the image after reconstruction (show in figure \ref{fig:treeEmbedOnceRestoration}) was 26.05dB.
The tamper rate, as was the case with the first image image (figure \ref{fig:lenaEmbedTwiceRestoration}), was increased from \textless 3\% to \textless 35\% with a similar reduction in quality experienced.
//TODO <Summarize all findings in an additional table>
\begin{center}
  \begin{tabular}{| l | l | l | l |}
    \hline
    \multicolumn{4}{|c|}{Experimental results summary} \\
    \hline
    Image & PSNR after embedding & PSNR after restoration & Tampering rate \\ \hline
    Figure \ref{fig:lenaEmbedOnceEmbedding} - \ref{fig:lenaEmbedOnceRestoration} (Embed once) & 34.92dB & 34.89dB & \textless 3\% \\ \hline
    Figure \ref{fig:lenaEmbedTwiceEmbedding} - \ref{fig:lenaEmbedTwiceRestoration} (Embed twice) & 34.92dB & 32.52 dB & \textless 35\% \\ \hline
    Figure \ref{fig:treeEmbedOnceEmbedding} - \ref{fig:treeEmbedOnceRestoration} (Embed once) & 27.94dB & 27.72dB & \textless 3\% \\ \hline
    Figure \ref{fig:treeEmbedTwiceEmbedding} - \ref{fig:treeEmbedTwiceRestoration} (Embed twice) & 27.93dB & 26.05dB & \textless 35\% \\ \hline
    \hline
  \end{tabular}
\end{center}

@@

\subsection{Summary}
The proposed method described used the BPCS steganaography algorithm, which is an adaptive embedding algorithm described in \ref{bpcsSteg}, and applied the algorithm to the self recovering image problem.
The proposed method was shown to provide good quality as well as a good tamper rate.
It was also shown that the quality of the image after embedding the image reference content as well as the image quality after reconstruction differs depending on the cover image chosen.
This was attributed to the adaptive embedding algorithm used due to the fact that the embedding differs depending on the cover image.
It was also mentioned that no compression was used on the image reference content that was embedded into the original image.
If compression is used it may result in higher quality of the image after reconstruction as well as the image after embedding due to fewer modifications that will be made during embedding.

>>>>

\title{Increasing quality of self-recovering images using BPCS Steganography}

@@

\noindent 
Many current self-recovering image methods exist, however few of the current methods provide good quality images after embedding as well as after reconstruction.
This is due to an inherent trade off between the amount of information that can be embedded into an image and the quality of the image after embedding.
This dissertation proposes a novel method for increasing the quality of the image after embedding as well as the quality of the image after reconstruction in the self-recovering image problem using the BPCS Steganography algorithm for content embedding as well as employing a fragile watermark for detecting image modifications.
Experimental results show that the proposed method achieves a good quality image after embedding as well as a good quality image after reconstruction when compared to some of the currently available self-recovering image methods.

@@

\subsection{Introduction}
During communication images can be modified either intentionally or unintentionally.  These modifications can be small changes that does not change the meaning of the image, or bigger changes, for example the removal of key aspects from the image. 
Various techniques exist for authenticating images of which self-recovering is one technique. 
Self-recovering techniques attempt to preserve image content by embedding the image within itself to enable the reconstruction of the image if the image was tampered with. Self-recovering techniques usually use steganography, a technology used to embed information in digital media \cite{johnson1998exploring}, to hide an approximation of the image within the image itself.  
If the image was tampered with, this embedded approximation can then be used to attempt to recover the original image.
 Self-recovering techniques also use fragile watermarks which is a technology commonly used in image authentication, to determine which areas of an image was tampered with.

\hspace{0pt} \\
Most self-recovering techniques have a trade-off concerning the quality of the image after embedding and the quality of the image after recovery.
If the one's quality is increased the other's quality will be reduced since an increase in the amount of embedded information could lead to a decrease in image quality and vice versa..

@@

\subsection{Problem statement}
\label{problemStatement}
The proposed method aims to answer the following questions:
\begin{itemize}
	\item Can a method be proposed that will achieve a high quality image after the self-embedding and restoration process as well as a reasonable tamper rate? How does such a system compare to currently available methods?
	\item Determining the quality of the resulting image should include, but are not limited to: 
	\begin{itemize}
		\item There should not be clear visual evidence that the image has been modified after embedding as well as after restoration
		\item The image should still carry it's original semantic meaning
	\end{itemize}
\end{itemize}
These questions will be addressed in chapters to follow.

\subsection{Research methodology}
Firstly a literature review was conducted to determine the currently available methods for self-recovering images. 
The proposed method was designed based on shortcomings that were found with currently available self-recovering methods. 
The proposed method was then implemented in the form of a Java software application in order to attempt to improve on the identified shortcomings. 
The results of the proposed method were then compared to the results of other relevant self-recovering image methods.

@@

\begin{itemize}
  \item Original Image: This refers to the original image without any modifications.
  \item Embedding: Embedding in the context of this dissertation refers to the process of storing digital information in an image in such a way that the original image is still distinguishable.
  \item Image after embedding: This refers to the the original image after the embedding process has been applied.
  \item Reconstruction: This refers to the process of using the content that was embedded into the image to repair the image after embedding if it has been modified.
  \item Image after reconstruction: This refers to the reconstructed version of the original image that was obtained by applying the reconstruction process to the image after embedding.
  \item LSB: This refers to the least significant bit in a binary string.
  \item MSB: This refers to the most significant bit/bits in a binary string.
  \item Section, Image section, Pixel section: This refers to a section of 8x8 pixels in an image.
  \item Bit plane: A bit plane is a 8x8 grid representation of a specific bit value for a specific pixel section. Each pixel has 24 bits which in total represents each colour, 8 bits per colour. These bits are the 8 binary bits that represent the numerical value of the colour. These 8 bits are separated from each other and are each stored in a 8x8 grid in the same location as that of the pixel in the pixel section. This process is repeated for each pixel in the pixel section and the result is 8, 8x8 grids of binary numbers. Each one of these 8 grids represent a single bit plane. 
  \item Block: This refers to an individual bit plane of a specific section of the image. There are 8 such blocks in each section since each pixel is represented by 8 bits for each colour.
\end{itemize}

\subsection{Chapter layout}
\begin{itemize}
\item Chapter 1 serves as an introduction to the concept of self-recovering images
\item Chapter 2 discusses self-recovering images and their different components
\item Chapter 3 discusses currently available self-recovering image schemes, comparing their strengths and weaknesses
\item Chapter 4 provides an overview of the system proposed in this dissertation
\item Chapter 5 discusses the results achieved during experimentation with the proposed system
\item Chapter 6 provides a conclusion, to determine if the questions stated in section \ref{problemStatement} have been answered
\end{itemize}

\subsection{Conclusion}
This chapter discussed the purpose behind self-recovering images as well as the aim of this research and the proposed method. 
The research methodology followed was described. A list of terminologies used in this dissertation was also given.
The next chapter will give a more detailed overview of all the components of self-recovering images.

@@

\hspace{0pt} \\
The PSNR of the image after reconstruction (show in figure \ref{fig:treeEmbedOnceRestoration}) was 26.05dB.
The tamper rate, as was the case with the first image image (figure \ref{fig:lenaEmbedTwiceRestoration}), was increased from \textless 3\% to \textless 35\% with a similar reduction in quality experienced.
\begin{center}
  \begin{tabular}{| l | l | l | l |}
    \hline
    \multicolumn{4}{|c|}{Experimental results summary} \\
    \hline
    Image & PSNR after embedding & PSNR after restoration & Tampering rate \\ \hline
    Figure \ref{fig:lenaEmbedOnceEmbedding} - \ref{fig:lenaEmbedOnceRestoration} (Embed once) & 34.92dB & 34.89dB & \textless 3\% \\ \hline
    Figure \ref{fig:lenaEmbedTwiceEmbedding} - \ref{fig:lenaEmbedTwiceRestoration} (Embed twice) & 34.92dB & 32.52 dB & \textless 35\% \\ \hline
    Figure \ref{fig:treeEmbedOnceEmbedding} - \ref{fig:treeEmbedOnceRestoration} (Embed once) & 27.94dB & 27.72dB & \textless 3\% \\ \hline
    Figure \ref{fig:treeEmbedTwiceEmbedding} - \ref{fig:treeEmbedTwiceRestoration} (Embed twice) & 27.93dB & 26.05dB & \textless 35\% \\ \hline
    \hline
  \end{tabular}
\end{center}

@@

\subsection{Summary}
The proposed method described used the BPCS steganaography algorithm, which is an adaptive embedding algorithm described in \ref{bpcsSteg}, and applied the algorithm to the self-recovering image problem.
The proposed method was shown to provide good quality as well as a good tamper rate.
It was also shown that the quality of the image after embedding the image reference content as well as the image quality after reconstruction differs depending on the cover image chosen.
This was attributed to the adaptive embedding algorithm used due to the fact that the embedding differs depending on the cover image.
It was also mentioned that no compression was used on the image reference content that was embedded into the original image.
If compression is used it may result in higher quality of the image after reconstruction as well as the image after embedding due to fewer modifications that will be made during embedding.
